---
title: "Issue analysis"
author: "Ntandoyenkosi and Evander"
date: "2024-03-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Packages

```{r}
library(tidyverse) # package for data manipulation and visualization.
library(janitor) # package for cleaning and tidying data.
# library("devtools")
# install_github("tmalsburg/saccades/saccades", 
#                dependencies=TRUE)
library(saccades) # package for analyzing eye movement data.
```



### Data

```{r}
subject_data_fin <- read_csv("subject_data/subject_data.csv")
```


```{r}
# Load external functions from "functions.R"
source("functions.R")

```


## First Issue

The screenshots of Subject 1 are the raw data file, the fixations (after running 
detect.fixations - please note I changed the smooth saccades to FALSE since I 
have low data frequency at 60hz, I also changed it to FALSE in your qmd file 
before running it), the summary stats and the Rscrip that I used for subject 1. 

For all the subject files with the x and y coordinates and time, which we use 
to detect fixations I removed the calibration screens from each of the subject 
folders. This means I manually removed screen 0 and the last screen of each 
participant (by implication this is screen 0 for all and screen 8 (treatment 
groups 1-4) and screen 10 (treatment groups 5-8).

The screenshots of the csv files with a focus on Subject 1 for reference and 
comparison.

You will see that even the fixations for subject 1 is not quite the same, and 
according to me both ran on detect.fixations, with the inputs and requirements 
(default of lambda is 6 in both).

Answer:



```{r}
subject_data_nw <- subject_data_fin %>%  # combine the data sets
  filter(!(screen_num %in% c(0, "0"))) %>%   # remove screen number 0
  mutate(treatment = case_when(
    subject %in% 1:33 ~ 1,
    subject %in% 34:66 ~ 2,
    subject %in% 67:96 ~ 3,
    subject %in% 97:128 ~ 4,
    subject %in% 129:160 ~ 5,
    subject %in% 161:193 ~ 6,
    subject %in% 194:226 ~ 7,
    subject %in% c(88, 114, 248) ~ 0,
    TRUE ~ 8 # Handle other cases if needed
  )) %>% 
  
  # Calculate groups based on treatment
  mutate(groups = case_when(treatment %in% 1:4 ~ "group 1",
                            T ~ "group 2")) %>%
  # mutate(groups = case_when(treatment %in% 1:4 ~ "group 1",
  #                         T ~ "group 2")) %>% 
  # Calculate selection based on groups and screen number conditions
  mutate(selection = case_when(groups == "group 1" & 
                                 screen_num == 8 ~ 1, #screen num 8 & Teatment 1-4
                               groups == "group 2" & 
                                 screen_num == 10 ~ 1, #screen num 10 & Teatment 5-8 
                               T ~ 0)) %>% 
  # Filter rows based on selection
  filter(selection != 1) %>% # remove selection
  # Remove the temporary selection column
  select(-selection) 
```



## Second Isuue


### Fixation by subject per screen number and number of views

We analyse fixations by subject per screen number per number of views.

```{r}
#' Calculate screen view statistics for each subject, screen, and visit 
#' combination.
screen_view_stats <- subject_data_fin %>%
  # Group by subject, screen_num, and num_visit_screen
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ process_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  mutate(stats = map(stats, ~ pivot_wider_and_clean(.x))) %>% 
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup() %>%
  # Rename columns with trial to screen_visit
  rename_with(~str_replace(., "trial", "screen_visit"), contains("trial")) %>% 
  # Select specific columns for the final output
  select(subject, screen_num, num_visit_screen, mean_no_of_fixations_per_screen_visit,
         mean_duration_of_fixations) %>% 
  # This column will help differentiate when we combina the fixations stats
  mutate(data = "subject_screen_visit")

# View output
#screen_view_stats %>% head(100) %>% view()

# Save screen_view_stats_output as csv file
write_csv(screen_view_stats,
          'subject_data/screen_view_stats_output_issue.csv')

```




### Issue 1

For instance subject 54. 
1. In the screen_stats_data.csv file there are 22 screens for this subject 
(out of a possible 23)  with screen number 3 being indicated as NA.  
Upon inspection of the fixations.csv file, the fixations on screen 3 visit 1 
(trial 3) are not included in fixations.csv file.

2. Screen 2 visit 4 (aka trial screen number 18) is shown in the fixations.csv 
file as NA - meaning no fixations on the screen, which would mean that this is 
the missing screen (22 out of 23)

#### Answer

1. In the screen_stats_data, subject 54 has only screen number 1-7. 
This dataset will not show the number of visits as it was grouped by only 
subject and screen 

By changing the smooth.saccades to False, this results to getting missing values

We only have the number of visits in the screen_view_stats dataset. This dataset,
indeed, shows the visit 1 for screen number 3 and is included.


2. Even though there is information for screen 2  and visit 4, no saccades were
detected for this instance. 


```{r}

screen_view_stats %>%
  filter(subject==54) %>% 
  count(screen_num)
  

screen_view_stats %>% 
  filter(subject == 54, screen_num ==3)
  # filter(subject %in% c(54, 72))


# Data for subject 54, screen number 2, visit ==4
subject_data_fin %>% 
  filter(subject == 54, screen_num ==2, num_visit_screen ==1)

# Fixations for subject 54, screen number 2, visit ==4
subject_data_fin %>% 
  filter(subject == 54, screen_num ==2, num_visit_screen ==4) %>%
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()


```

### Issue 2

Then I have a different scenario, looking at subject 72:
According to the screen_stats_data file subject 72 has 26 screens . According to 
the fixations file, screen 5 visit 5 (aka trial screen number 26) has no 
fixations and is indicated as NA

The problem however is, upon further inspection of the fixations file, I find 
that the fixations for screen 2 visit 3 (aka trial 22) are not listed in the 
fixations file, but the screen_stats_data file shows the summary statistics 
for 3 x  screen 2 visits, which is the correct number of visits to screen 2 as 
per the raw data. Now my question is, how can we have summary statistics for a 
screen when there are no fixations listed for that screen in the fixations file. 
The screen_view_stats_output.csv file shows the same summary statistics and 
confirms that screen 2 visit 3 has summary statistics.


### Answer

From our analysis subject 72, screen number 5 and visit 5 shows that no fixations 
can be obtained from such instance as there are no saccades that can be obtained

Upon investigating further there are possible reasons for not getting any fixations

1. Discontinuity in the data
2. Poor data


Source:

1. https://stackoverflow.com/questions/54997537/problems-in-the-analysis-of-gaze-data-r-package-saccades#:~:text=I%20suspect%20that%20there%20are,parse%20into%20saccades%20and%20fixations.&text=If%20ggplot2%20isn't%20installed,packages('ggplot2')%20first.


We expolored the 2 possibilities. There is no discontinuity in the data as shown
by the following graph. This leaves us with the poor data problem.

For subject 72, screen number 2 and visit 3, fixations can be obtained as below


```{r}
# Data for subject 72, screen number 5, visit ==5
subject_72_5_5 <- subject_data_fin %>% 
  filter(subject == 72, screen_num ==5, num_visit_screen ==5) %>% 
  select(time, x, y, trial)

# Fixations for subject 72, screen number 5, visit ==5
subject_72_5_5_fix <- subject_data_fin %>% 
  filter(subject == 72, screen_num ==5, num_visit_screen ==5) %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()


# Fixations 
subject_72_5_5_fix


# Plot to see discontinuities in the data
ggplot(data = subject_72_5_5,
       aes(x = time)) + # define time series x axis
geom_line(aes(y = x, colour = 'X')) +  # plot horizontal data
geom_line(aes(y = y, colour = 'Y')) + # plot vertical time series too
scale_color_manual(name = "Coordinates", values = c("X" = "darkblue", 
                                                 "Y" = "darkred")) +
labs(x = "Time", y = "X and Y lines",
     title = "X, Y coordinates by Time") +
  theme_bw()

# Fixations for subject 72, screen number 5, visit ==5
subject_data_fin %>% 
  filter(subject == 72, screen_num ==2, num_visit_screen ==3) %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()
  
  
```



Use the following code to get the fixations for all instances:

A total of 428 instances have missing fixations. We can identify 5097 fixations
for different instances of (subject, screen number and visits)

```{r}

subject_screen_visit_fixations <- subject_data_fin %>%
  select(-value) %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()


subject_screen_visit_fixations

subject_screen_visit_fixations %>% 
  filter(!is.na(x)) %>% 
  group_by(subject, screen_num, num_visit_screen) %>% 
  count()

subject_screen_visit_fixations %>% 
  filter(is.na(x)) %>% 
  group_by(subject, screen_num, num_visit_screen) %>% 
  count()

write_csv(subject_screen_visit_fixations,
          "subject_data/subject_screen_visit_fixations.csv")

```

