---
title: "eye_study_analysis_report"
author: "UI Labs"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Background

This report details how the fixation analysis was done by:

1.  Following a recommended folder structure
2.  Applying some functions inline with the Don't Repeat Yourself (DRY) principle
3.  Combining the subject .csv files into a single csv
4.  . Analysing fixations per:
    -   Subject

    -   Subject per screen number

    -   Subject per screen number per number of visits

## Analysis

### Folder structure

In order to use this code, we need to create an eye study project in R, this helps us to easily navigate and use the code among ourselves irrespective of where (downloads/documents/desktop etc) we saved our datasets. By following this step, there won't be a need to do perform **'*setwd()*'**.

Please follow the following directions in creating a project folder:

1.  Click on File (top left corner)
2.  Click on New project
3.  Click on New Directory
4.  Click on New project
5.  Give a directory name of your choice, put underscore where there are spaces (we used eye_study)
6.  The subdirectory is a folder that you want to store this project in. For example you can create a folder in your documents called ***'playground'*** most people do this, it will be a folder where you store all future projects in.
7.  Once done, you will need to past in the data folders you shared with us in the new created project folder
8.  Create a new folder called subject_data, this is where we will save the ***.csv files*** we will be creating.

We believe by using this folder strategy, the code becomes more reproducible and can be easily shared among users.

### Packages

The following packages were used in the analysis. To install the SACCADES package, uncomment and load the devtools package and run the installation of the saccades from github.

```{r}
# Load necessary R packages for data analysis 

library(tidyverse) # package for data manipulation and visualization.
library(janitor) # package for cleaning and tidying data.
# library("devtools")
# install_github("tmalsburg/saccades/saccades", 
#                dependencies=TRUE)
library(saccades) # package for analyzing eye movement data.

```

### Functions

Inline with the Dont Repeat Yourself (DRY) coding principle, we developed some functions to help with the data wrangling and fixation analysis. We have saved the functions in an R script called functions.

```{r}
# Load external functions from "functions.R"
source("functions.R")

```

### Data extraction

The following code, gets all .csv files and combines them into one csv file. [**This will take sometime to run**]{.underline}

```{r}
# List all folders in the specified directory containing subject data.
subject_folders <- list.files("data/PhD data shared/Mapped data per subject (per trial)") # List all folders

# Create full paths for each subject folder.
subject_folders <- str_c("data/PhD data shared/Mapped data per subject (per trial)", '/', subject_folders)# Combine the folder structure name and the subject folders

# Convert list of folders to tibble and read CSV files
subject_data <- subject_folders %>% 
  as_tibble() %>% 
  mutate(data = map(.x = value, .f = ~csv_reader(.x))) %>% 
  unnest(data) # read the csv files

# Filter rows with missing values in specific columns
subject_data_1 <- subject_data %>% 
  filter(is.na(`Subject\tTreatment\tscreen_num\tnum_visit_screen\tx\ty\ttime\ttrial`)) %>% 
  select(value:trial) %>% 
  clean_names() 

# Filter out rows with missing values in specific columns
subject_data_2 <- subject_data %>% 
  filter(!is.na(`Subject\tTreatment\tscreen_num\tnum_visit_screen\tx\ty\ttime\ttrial`)) %>% 
  # Select relevant columns
  select(value, `Subject\tTreatment\tscreen_num\tnum_visit_screen\tx\ty\ttime\ttrial`) %>% 
  # Separate the combined column into individual columns
  separate(col = `Subject\tTreatment\tscreen_num\tnum_visit_screen\tx\ty\ttime\ttrial`,
           into = c("subject", "treatment",	"screen_num",	"num_visit_screen",	
                    "x",	"y",	"time",	"trial"),
           sep = '\t') %>% 
  # Convert columns from character to double
  mutate(across(.cols = subject:trial, .fns = ~as.double(.x)))


# Combine the two data sets vertically
subject_data_fin <- bind_rows(subject_data_1, subject_data_2)# %>%  # combine the data sets
  # filter(screen_num != 0, screen_num == "0") # remove screen number 0


# Save as csv
write_csv(subject_data_fin, 'subject_data/subject_data.csv') # save the csv


# Load/Read the csv data
# subject_data_fin <- read_csv("subject_data/subject_data.csv") # Yo can load the data later

```



### Testing

```{r}

# check missing values - x, y, time & trial

subject_data_stats <- subject_data_fin %>% 
  filter(!(screen_num %in% c(0, "0"))) %>% 
  group_by(subject) %>% 
  summarise(
    n = n(),
    n_trials = n_distinct(trial),
    n_screens = n_distinct(screen_num),
    n_visits = n_distinct(num_visit_screen),
    missing_x = sum(is.na(x)),
    missing_y = sum(is.na(y)),
    missing_time = sum(is.na(time)),
    missing_trial = sum(is.na(trial))
  ) 

subject_data_stats %>% 
  mutate(missing_tot = missing_x+ missing_y + missing_time + missing_trial) %>% 
  filter(missing_tot >0)
  # filter(subject %in% c(54, 72))

subject_data_fin %>% 
  filter(!screen_num %in% c(0, "0")) %>% 
  filter(subject %in% c(54, 72))

```





### Fixation by subject and screen number

Using the combined .csv file, we analyse fixations per subject per screen number. We then filtered screen number 1-7 for treatment 1-4 and screen number 1-9 for treatment 5-8.

```{r}
screen_stats <- subject_data_fin %>%
  # Group data by subject and screen number
  group_by(subject, screen_num) %>% 
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate screen-level statistics using custom function process_possibly
  mutate(stats = map(data, ~ process_possibly(.x))) %>% 
  mutate(stats = map(stats, ~ pivot_wider_and_clean(.x))) %>% 
  # Unnest the statistics
  unnest(stats) %>% 
  # Remove the original nested data
  select(-data) %>% 
  # Remove grouping
  ungroup() %>%
  # Rename columns related to trials to be related to screens
  rename_with(~str_replace(., "trial", "screen"), contains("trial")) %>%
  # Select relevant columns
  select(subject, screen_num, mean_no_of_fixations_per_screen,
         mean_duration_of_fixations)
  
#View the ouput
#screen_stats %>% head(100) %>% view()



screen_stats_output <- screen_stats %>% 
  # Generate treatment assignments 
  mutate(treatment = case_when(
    subject %in% 1:33 ~ 1,
    subject %in% 34:66 ~ 2,
    subject %in% 67:96 ~ 3,
    subject %in% 97:128 ~ 4,
    subject %in% 129:160 ~ 5,
    subject %in% 161:193 ~ 6,
    subject %in% 194:22  ~ 7,
    subject %in% c(88, 114, 248) ~ 0,
    TRUE ~ 8 # Handle other cases if needed
  )) %>% 
  
  # Calculate groups based on treatment
  mutate(groups = case_when(treatment %in% 1:4 ~ "group 1",
                            T ~ "group 2")) %>%
  # mutate(groups = case_when(treatment %in% 1:4 ~ "group 1",
  #                         T ~ "group 2")) %>% 
  # Calculate selection based on groups and screen number conditions
  mutate(selection = case_when(groups == "group 1" & screen_num %in% 1:7 ~ 1,
                               groups == "group 2" & screen_num %in% 1:9 ~ 1, 
                               T ~ 0)) %>% 
  # Filter rows based on selection
  filter(selection == 1) %>%
  # Remove the temporary selection column
  select(-selection) %>% 
  # This column will help differentiate when we combina the fixations stats
  mutate(data = "subject_screen")
  

# Save screen_stats as csv file
write_csv(screen_stats_output,
          'subject_data/screen_stats_data.csv')

# View the output
#screen_stats_output %>% head(100) %>% view()
```







### Fixation by subject per screen number and number of views

We analyse fixations by subject per screen number per number of views.

```{r}
#' Calculate screen view statistics for each subject, screen, and visit 
#' combination.
screen_view_stats <- subject_data_fin %>%
  # Group by subject, screen_num, and num_visit_screen
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ process_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  mutate(stats = map(stats, ~ pivot_wider_and_clean(.x))) %>% 
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup() %>%
  # Rename columns with trial to screen_visit
  rename_with(~str_replace(., "trial", "screen_visit"), contains("trial")) %>% 
  # Select specific columns for the final output
  select(subject, screen_num, num_visit_screen, mean_no_of_fixations_per_screen_visit,
         mean_duration_of_fixations) %>% 
  # This column will help differentiate when we combina the fixations stats
  mutate(data = "subject_screen_visit")

# View output
#screen_view_stats %>% head(100) %>% view()

# Save screen_view_stats_output as csv file
write_csv(screen_view_stats,
          'subject_data/screen_view_stats_output.csv')

```

### Fixation by subject

We then analyse the fixations per subject and save the csv file

```{r}
#' Compute statistics for each subject based on the provided data.
subject_stats <- subject_data_fin %>%
  # Group data by subject
  group_by(subject) %>% 
  nest(data = c(treatment, screen_num, num_visit_screen, x, y, time, trial)) %>% 
  # Compute statistics for each group
  mutate(stats = map(data, ~ process_possibly(.x))) %>% 
  # Pivot and clean the computed statistics
  mutate(stats = map(stats, ~ pivot_wider_and_clean(.x))) %>% 
  unnest(stats) %>% 
  # Select relevant columns and remove unnecessary ones
  select(-c(data, value)) %>% 
  ungroup() %>%
  # Rename columns with subject-specific information
  rename_with(~str_replace(., "trial", "subject"), 
              contains("trial")) %>%
  # Select final set of columns
  select(subject, 
         mean_duration_of_subjects,
         mean_duration_of_fixations,
         mean_no_of_fixations_per_subject)  %>% 
  # This column will help differentiate when we combina the fixations stats
  mutate(data = "subject")
  
# View the data
# subject_stats %>% 
#   head(100) %>% 
#   view()

# Save as csv
write_csv(subject_stats,
          'subject_data/subject_stats.csv')

```

### Combining the data

We then combine the fixation stats datasets into one single csv as follows

```{r}
fixation_stats <- bind_rows(subject_stats,
                            screen_stats_output,
                            screen_view_stats)

# Save as csv
write_csv(fixation_stats,
          'subject_data/fixation_stats.csv')
```
