---
title: "Fixations analysis"
author: "Ntando and Evander"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Purpose

```{r}
library(tidyverse) # package for data manipulation and visualization.

library(janitor) # package for cleaning and tidying data.

# library("devtools")

# install_github("tmalsburg/saccades/saccades",

# dependencies=TRUE)

library(saccades) # package for analyzing eye movement data.
```

### Data

```{r}
subject_data_fin <- read_csv("subject_data/subject_data.csv") %>% 
  filter(!(screen_num %in% c(0, "0"))) %>%    # remove screen number 0
  select(-value) # remove folder roots
```

### Functions

```{r}
# Load external functions from "functions.R"
source("functions.R")
```

### Fixations for subject 72

From our analysis subject 72, screen number 5 and visit 5 shows that no fixations

can be obtained from such instance as there are no saccades that can be obtained

Upon investigating further there are possible reasons for not getting any fixations

1\. Discontinuity in the data

2\. Poor data

Source:

1\. https://stackoverflow.com/questions/54997537/problems-in-the-analysis-of-gaze-data-r-package-saccades#:\~:text=I%20suspect%20that%20there%20are,parse%20into%20saccades%20and%20fixations.&text=If%20ggplot2%20isn't%20installed,packages('ggplot2')%20first.

We expolored the 2 possibilities. There is no discontinuity in the data as shown

by the following graph. This leaves us with the poor data problem.

For subject 72, screen number 2 and visit 3, fixations can be obtained as below

```{r}
subject_72_5_5 <- subject_data_fin %>% 
  filter(subject == 72, screen_num ==5, num_visit_screen ==5) %>% 
  select(time, x, y, trial)

# Fixations for subject 72, screen number 5, visit ==5
subject_72_5_5_fix <- subject_data_fin %>% 
  filter(subject == 72, screen_num ==5, num_visit_screen ==5) %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()


# Fixations 
subject_72_5_5_fix


# Plot to see discontinuities in the data
ggplot(data = subject_72_5_5,
       aes(x = time)) + # define time series x axis
geom_line(aes(y = x, colour = 'X')) +  # plot horizontal data
geom_line(aes(y = y, colour = 'Y')) + # plot vertical time series too
scale_color_manual(name = "Coordinates", values = c("X" = "darkblue", 
                                                 "Y" = "darkred")) +
labs(x = "Time", y = "X and Y lines",
     title = "X, Y coordinates by Time") +
  theme_bw()


```

There is no discontinuity in the data meaning that the missing data points could be a result of poor data

## Fixations: Subject

```{r}

subject_fixations <- subject_data_fin %>%
  # group_by(subject, screen_num) %>%
  group_by(subject) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()

# Save the fixations data
write_csv(subject_fixations,
          "subject_data/subject_fixations.csv")

# Get the first 6 rows output
subject_fixations %>% 
  head()

# Count the events (fixations, blinks etc) 
subject_fixations_count <- subject_fixations %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject) %>%
  count(event) %>% 
  mutate(percentage = scales::percent(n/sum(n)))

# Take a look at the first 6 rows
subject_fixations_count %>% 
  head()

# Save this data
write_csv(subject_fixations_count,
          "output/subject_fixations_count.csv")

# Total rows of info
total_rows <- nrow(subject_fixations %>% 
  # filter(!is.na(x)) %>% 
  group_by(subject) %>% 
  count()
  )

# Non missing information
total_non_missing_rows <- nrow(subject_fixations %>% 
  filter(!is.na(x)) %>%
  group_by(subject) %>% 
  count()
  )

#percentage
non_missing_perc <- round((total_non_missing_rows/total_rows)*100,2)
missing_perc <- round(100 - as.numeric(non_missing_perc),2)

# Percentage report
glue::glue("The total percentage of data that is missing is {missing_perc}%, meaning we have a total of {non_missing_perc}% non missing data points")
```

## Fixations: Subject and Screen number

```{r}
subject_screen_fixations <- subject_data_fin %>%
  # group_by(subject, screen_num) %>%
  group_by(subject, screen_num) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()

# Save the fixations data
write_csv(subject_screen_fixations,
          "subject_data/subject_screen_fixations.csv")

# Get the first 6 rows output
subject_screen_fixations %>% 
  head()

# Count the events (fixations, blinks etc) 
subject_screen_fixations_count <- subject_screen_fixations %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num) %>%
  count(event) %>% 
  mutate(percentage = scales::percent(n/sum(n)))

# Take a look at the first 6 rows
subject_screen_fixations_count %>% 
  head()

# Save this data
write_csv(subject_screen_fixations_count,
          "output/subject_screen_fixations_count.csv")

# Total rows of info
total_rows_2 <- nrow(subject_screen_fixations %>% 
  # filter(!is.na(x)) %>% 
  group_by(subject, screen_num) %>% 
  count()
  )

# Non missing information
total_non_missing_rows_2 <- nrow(subject_screen_fixations %>% 
  filter(!is.na(x)) %>%
  group_by(subject, screen_num) %>% 
  count()
  )

#percentage
non_missing_perc_2 <- round((total_non_missing_rows_2/total_rows_2)*100,2)
missing_perc_2 <- round(100 - as.numeric(non_missing_perc_2),2)

# Percentage report
glue::glue("The total percentage of data that is missing is {missing_perc_2}%, meaning we have a total of {non_missing_perc_2}% non missing data points")
```

## Fixations: Subject, Screen and Visit

```{r}
subject_screen_visit_fixations <- subject_data_fin %>%
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  # Nest data for each group
  nest(data = c(treatment, x, y, time, trial)) %>% 
  # Calculate statistics for nested data using process_possibly
  mutate(stats = map(data, ~ fixations_possibly(.x))) %>% 
  # Pivot and clean the resulting statistics
  # Unnest the statistics
  unnest(stats) %>% 
  # Select relevant columns and remove the nested data column
  select(-data) %>% 
  # Ungroup the data
  ungroup()

# Save the fixations data
write_csv(subject_screen_visit_fixations,
          "subject_data/subject_screen_visit_fixations.csv")

# Get the first 6 rows output
subject_screen_visit_fixations %>% 
  head()

# Count the events (fixations, blinks etc) 
subject_screen_visit_fixations_count <- subject_screen_visit_fixations %>% 
  # group_by(subject, screen_num, num_visit_screen) %>%
  group_by(subject, screen_num, num_visit_screen) %>%
  count(event) %>% 
  mutate(percentage = scales::percent(n/sum(n)))

# Take a look at the first 6 rows
subject_screen_visit_fixations_count %>% 
  head()

# Save this data
write_csv(subject_screen_visit_fixations_count,
          "output/subject_screen_visit_fixations_count.csv")

# Total rows of info
total_rows_3 <- nrow(subject_screen_visit_fixations %>% 
  # filter(!is.na(x)) %>% 
  group_by(subject, screen_num, num_visit_screen) %>% 
  count()
  )

# Non missing information
total_non_missing_rows_3 <- nrow(subject_screen_visit_fixations %>% 
  filter(!is.na(x)) %>%
  group_by(subject, screen_num, num_visit_screen) %>% 
  count()
  )

#percentage
non_missing_perc_3 <- round((total_non_missing_rows_3/total_rows_3)*100,2)
missing_perc_3 <- 100 - as.numeric(non_missing_perc_3)

# Percentage report
glue::glue("The total percentage of data that is missing is {missing_perc_3}%, meaning we have a total of {non_missing_perc_3}% non missing data points")



```
